Insights Into Computational Reproducibility in Plant Pathology and a Way Forward

### 2020-12-07 analysis/paper/paper.Rmd
Start hopefully final edits on paper to send to coauathors soon

---
title: "How Can We Open Plant Pathology?"
author:
  - Adam H. Sparks:
      email: Adam.Sparks@usq.edu.au
      institute: USQ
      correspondence: true
  - Emerson Del Ponte:
      institute: UFV
  - Kaique dos Santos Alves:
      institute: UFV
  - Zachary Foster: 
      institute: OSU
  - Niklaus J Grünwald: 
      institute: ARS
institute:
  - USQ: University of Southern Queensland, Centre for Crop Health, Toowoomba, Queensland, Australia
  - UFV: Departmento de Fitopatologia, Universidade Federal de Viçosa, Brazil
  - OSU: Department of Botany and Plant Pathology, Oregon State University, Corvallis, OR, USA
  - ARS: Horticultural Crops Research Unit, USDA Agricultural Research Service, Corvallis, OR, USA
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  officedown::rdocx_document:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
      pandoc_args:
        - '--lua-filter=scholarly-metadata.lua'
        - '--lua-filter=author-info-blocks.lua'
        - '--lua-filter=pagebreak.lua'
      tables:
        layout: autofit
        width: 1.0
        caption:
          style: Table Caption
          pre: 'Table '
          sep: ': '
        conditional:
          first_row: true
          first_column: false
          last_row: false
          last_column: false
          no_hband: false
          no_vband: true
bibliography: references.bib
csl: "../templates/phytopathology.csl" # Insert path for the bib-style
abstract: |
  Open research practices have been highlighted extensively during the last ten years in many fields of study as essential standards needed to promote transparency and reproducibility of scientific results. In fact, scientific claims can only be evaluated based on how protocols, materials, equipment and methods were described; data were collected and prepared; and analyses conducted. Openly sharing protocols, data and computational code are central for current scholarly dissemination and communication, but in many fields, including plant pathology, adoption of these practices has been slow. We randomly selected 300 articles published from 2012 to 2018 across 21 journals representative of the pathology discipline, and assigned them scores reflecting their openness and reproducibility. We found that most of the articles were not very open, failing to share data or code in a meaningful way. We propose that using open source tools for producing open and reproducible work and analysis is advantageous, benefiting not just readers, but the authors as well and provide ideas and tools to promote open, reproducible research practices among plant pathologists.
keywords: |
  open science; open data; reproducible research; FAIR data
highlights: |
  Open research practices are highlighted extensively in the broader scientific community. Our survey of 300 randomly selected articles from 2012 to 2018 indicates that we have work to do in this area as plant pathologists with most articles scoring low in openness and reproducibility.
---

@@

However, while these definitions are clear and well accepted in physical sciences, in biological sciences it is not always possible to use identical test material or perhaps the the time or resources are not available for full reproduction, _e.g._, field trials that span years and locations or complex glasshouse experiments.
Therefore we will follow Peng's [-@Peng2009] definition that provides clear guidelines a minimum standard of ``reproducible research'':

> "The replication of scientific findings using independent investigators, methods, data, equipment, and protocols has long been, and will continue to be, the standard by which scientific claims are evaluated. However, in many fields of study there are examples of scientific investigations that cannot be fully replicated because of a lack of time or resources. In such a situation, there is a need for a minimum standard that can fill the void between full replication and nothing. One candidate for this minimum standard is ``reproducible research'', which requires that data sets and computer code be made available to others for verifying published results and conducting alternative analyses". Peng, R. [-@Peng2009]. _Reproducible research and Biostatistics_. Biostatistics, 10 (3): 405-408.

### Open Science

Open science has become highlighted lately with many donors expecting data to be available [@government_of_canada_2016; @vannoorden2017; @ARC2018] and other scientists interested in sharing and collaborating more widely [@Wald2010].

### Methods

# Status in Plant Pathology

@@

In order for us to determine where we are as a discipline with regards to sharing openly and reproducible papers and data, we randomly selected 300 articles published from 2012 to 2018 across 21 journals (Table \@ref(tab:journals)) representative of the pathology discipline, and assigned them scores reflecting their openness and reproducibility.
The journals represented both fundamental and applied areas as well as pathogen or other group specific publications.
Three hundred articles were randomly selected from issues published from 2012 to 2018.
A list of randomly selected pages was assigned to a randomized list of the 21 journals where the page number fell within an article for the given journal.
In cases where an article was not suitable, _e.g._, a review or otherwise not related to plant pathology, the next article was selected until a suitable article was found.
Notes regarding the selection of articles can be found in the file, XXXX, available in this paper's repository.
The pages list was numbered from page one and went to page 150.
This was done since some journals restart their numbering with each issue and also ensures that the journal is more likely to have a page number corresponding to the randomly generated value.
This also assumes that there is no effect or bias on reproducibility based on the time of year that an article was published, since most journals start with page number one at the beginning of the year.
The list of journals was saved from Google's sheets as a comma separated value (CSV) file for later use and distribution in a research compendium.

@@

Each journals was classified as to  whether they were completely open (TRUE), behind a pay wall (FALSE) or a combination (BOTH).
The five-year impact factor for 2018 for each journal was retrieved from InCites Journal Citation Reports, Clarivate Analytics and entered in a separate sheet in the Google sheets file.
This was downloaded, saved as a CSV file and is left-joined with the article notes using `dplyr::left_join()` for analysis when the `import_notes()` command is executed.

@@

    * 2 - Uses proprietary software that most individuals can afford (_e.g._, Excel) or SAS that is now available free for use with a univerisity login

@@

We recognise that the scoring methods were developed by the auhors specifically for this exercise and are not entirely subjective.
We have based our scoring decisions on our own experiences in the field of plant pathology.
While many software packages may be widely used or "freely" available if one has an institutional licence, these types of software often hinder partners and interested parties that do not have these luxuries.
Thus, emphasis was given to using open source alternatives where possible over paid software packages.
The view of the authors is one that this is something to be encouraged as it allows the broadest possible audience to benefit from the work at the least cost.

@@

Forty-eight articles scored 0&nbsp;% out of a possible 100&nbsp;% for reproducibility.
The mean value was 31.2&nbsp;% and the median value was 33.3&nbsp;%.
One article scored 100&nbsp;%, or was classed as fully reproducible.

@@

Most articles did not make any computational methods available in any fashion with two, 0.007&nbsp;%, classing as "3", which was the highest score available.
Two hundred eighty seven, 96&nbsp;%, were classed as "0" and 9 articles, 0.03&nbsp;%, appeared to not use any computational methods.

@@

A general workflow for producing academic research involves clearly defining a research question, obtaining data for testing the hypothesis, summarizing/analyzing and presenting data and results, and writing the manuscript.
Here we defined three levels of reproducibility which are also related with the evolution of computational methods and reproducible practices (Figure \@ref(fig:workflow-dia)).

>>>>

---
title: "Insights Into Computational Reproducibility in Plant Pathology and a Way Forward"
author:
  - Adam H. Sparks:
      email: Adam.Sparks@usq.edu.au
      institute: USQ
      correspondence: true
  - Emerson Del Ponte:
      institute: UFV
  - Kaique dos Santos Alves:
      institute: UFV
  - Zachary Foster: 
      institute: OSU
  - Niklaus J Grünwald: 
      institute: ARS
institute:
  - USQ: University of Southern Queensland, Centre for Crop Health, Toowoomba, Queensland, Australia
  - UFV: Departmento de Fitopatologia, Universidade Federal de Viçosa, Brazil
  - OSU: Department of Botany and Plant Pathology, Oregon State University, Corvallis, OR, USA
  - ARS: Horticultural Crops Research Unit, USDA Agricultural Research Service, Corvallis, OR, USA
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  officedown::rdocx_document:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
      pandoc_args:
        - '--lua-filter=scholarly-metadata.lua'
        - '--lua-filter=author-info-blocks.lua'
        - '--lua-filter=pagebreak.lua'
      tables:
        layout: autofit
        width: 1.0
        caption:
          style: Table Caption
          pre: 'Table '
          sep: ': '
        conditional:
          first_row: true
          first_column: false
          last_row: false
          last_column: false
          no_hband: false
          no_vband: true
bibliography: references.bib
csl: "../templates/phytopathology.csl" # Insert path for the bib-style
abstract: |
  Open research practices have been highlighted extensively during the last ten years in many fields of study as essential standards needed to promote transparency and reproducibility of scientific results. In fact, scientific claims can only be evaluated based on how protocols, materials, equipment and methods were described; data were collected and prepared; and analyses conducted. Openly sharing protocols, data and computational code are central for current scholarly dissemination and communication, but in many fields, including plant pathology, adoption of these practices has been slow. We randomly selected 300 articles published from 2012 to 2018 across 21 journals representative of the pathology discipline, and assigned them scores reflecting their openness and reproducibility. We found that most of the articles were not very open, failing to share data or code in a meaningful way. We propose that using open source tools for producing open and reproducible work and analysis is advantageous, benefiting not just readers, but the authors as well and provide ideas and tools to promote open, reproducible research practices among plant pathologists.
keywords: |
  open science; open data; reproducible research; FAIR data
highlights: |
  Open research practices are highlighted extensively in the broader scientific community. Our survey of 300 randomly selected articles from 2012 to 2018 indicates that we have work to do in this area as plant pathologists with most articles scoring low in openness and reproducibility.
---

@@

However, while these definitions are clear and well accepted in physical sciences, in biological sciences it is not always possible to use identical test material or perhaps the the time or resources are not available for full reproduction, _e.g._, field trials that span years and locations or complex glasshouse experiments.
Therefore we will follow Peng's [-@Peng2009] definition that provides clear guidelines a minimum standard of "reproducible research":

> "The replication of scientific findings using independent investigators, methods, data, equipment, and protocols has long been, and will continue to be, the standard by which scientific claims are evaluated. However, in many fields of study there are examples of scientific investigations that cannot be fully replicated because of a lack of time or resources. In such a situation, there is a need for a minimum standard that can fill the void between full replication and nothing. One candidate for this minimum standard is 'reproducible research', which requires that data sets and computer code be made available to others for verifying published results and conducting alternative analyses". Peng, R. [-@Peng2009]. _Reproducible research and Biostatistics_. Biostatistics, 10 (3): 405-408.

Therefore our definition of reproducibility will be that the computer code and data are made freely available to others for verification and conducting alternate analyses or for use in instructional purposes and that software used are also easily obtainable and preferably open source to avoid licensing or other issues related to accessibility for end-users related to costs or non-standard file formats, etc.

## On Open Science and Reproducibility

Open science has become highlighted lately with many donors expecting data to be available [@government_of_canada_2016; @vannoorden2017; @ARC2018] and other scientists interested in sharing and collaborating more widely [@Wald2010].
While open science and reproducibility are not the same, they are linked.
By proactively practising open science, the work becomes more reproducible through the availability of data and code.

# Methods

We conducted a survey of a broad selection of articles to represent a broad swathe of publications in our discipline to evaluate the status of reproducibility and open science in our field.
The journals represented both fundamental and applied areas as well as pathogen or other group specific publications.

@@

We randomly selected 300 articles published from 2012 to 2018 across 21 journals (Table \@ref(tab:journals)) and assigned them scores reflecting their openness and reproducibility.
A list of randomly selected pages was assigned to a randomized list of the 21 journals where the page number fell within an article for the given journal.
In cases where an article was not suitable, _e.g._, a review or otherwise not related to plant pathology, the next article was selected until a suitable article was found.
Notes regarding the selection of articles can be found in the file, XXXX, available in this paper's repository.
The pages list was numbered from page one and went to page 150.
This was done since some journals restart their numbering with each issue and also ensures that the journal is more likely to have a page number corresponding to the randomly generated value.
This also assumes that there is no effect or bias on reproducibility based on the time of year that an article was published, since most journals start with page number one at the beginning of the year.
The list of journals was saved from Google's sheets as a comma separated value (CSV) file for later use and distribution in a research compendium.

@@

Each journals was classified as to  whether they were completely open (TRUE), behind a pay wall (FALSE) or a combination (BOTH).
The five-year impact factor for 2018 for each journal was retrieved from InCites Journal Citation Reports, Clarivate Analytics and entered in a separate sheet in the Google sheets file.
This was downloaded, saved as a CSV file and is left-joined with the article notes using a `left_join()` from the 'dplyr' package [@Wickham2020] for analysis when the `import_notes()` command is executed.

@@

    * 2 - Uses proprietary software that most individuals can afford (_e.g._, Excel) or SAS that is now available free for use with a university login

@@

We recognise that the scoring methods were developed by the authors specifically for this exercise and are not entirely subjective.
We have based our scoring decisions on our own experiences in the field of plant pathology.
While many software packages may be widely used or "freely" available if one has an institutional licence, these types of software often hinder partners and interested parties that do not have these luxuries.
Thus, emphasis was given to using open source alternatives where possible over paid software packages.
The view of the authors is one that this is something to be encouraged as it allows the broadest possible audience to benefit from the work at the least cost.

@@

Forty-eight articles scored 0% out of a possible 100% for reproducibility.
The mean value was 31.2% and the median value was 33.3%.
One article scored 100%, or was classed as fully reproducible.

@@

Most articles did not make any computational methods available in any fashion with two, 0.007%, classing as "3", which was the highest score available.
Two hundred eighty seven, 96%, were classed as "0" and 9 articles, 0.03%, appeared to not use any computational methods.

@@

A general work flow for producing academic research involves clearly defining a research question, obtaining data for testing the hypothesis, summarizing/analyzing and presenting data and results, and writing the manuscript.
Here we defined three levels of reproducibility which are also related with the evolution of computational methods and reproducible practices (Figure \@ref(fig:workflow-dia)).
